
---
title: "Understanding Data"
author: "Andy Areitio"
output: html_document
---


<br>

The purpose of this exercise is to understand our DATA.

#1 --- *ANDY*: this in case that we need to donwload data sometime!
### Exercise Data

We download daily prices (open, high, low, close, and adjusted close) and volume data of publicly traded companies and markets from the web (e.g. Yahoo! or Google, etc). This is done by sourcing the file data.R as well as some helper functions in herpersSet1.R which also installs a number of R libraries (hence the first time you run this code you will see a lot of red color text indicating the *download* and *installation* process):

```{r eval = TRUE, echo=TRUE, error = FALSE, warning=FALSE,message=FALSE,results='asis'}
#this files doesnt exist... it is just to remind us!
source("helpersSet1.R")
source("dataSet1.R")
```


#2 --- *ANDY*: this is to come up with basic understanding: num colums, rowns, means,...


###  Part IV: Read/Write .CSV files

Finally, one can read and write data in .CSV files. For example, we can save the first 20 days of data for S&P and Apple in a file using the command:

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
write.csv(StockReturns[1:20,c("SPY","AAPL")], file = "twentydays.csv", row.names = TRUE, col.names = TRUE) 
```

Do not get surpsised if you see the csv file in your directories suddenly! You can then read the data from the csv file using the read.csv command. For example, this will load the data from the csv file and save it in a new variable that now is called "myData": 

```{r eval = TRUE, echo=TRUE, comment=NA, warning=FALSE, message=FALSE,results='asis'}
myData <- read.csv(file = "NYPD_Motor_Vehicle_Collisions_2016.csv", header = TRUE, sep=";")
class(myData)
dim(myData)
nrow(myData)
ncol(myData)
myData[1:4,]
head(myData,5)
tail(myData,5)
round(100*sum(myData[,10]),1)
round(100*sum(myData[,1]),1)
round(100*sum(myData[,5]),1)
round(100*sum(myData[,5]),1)
round(100*mean(myData[,1]),1)
round(100*mean(myData[,5]),1)
round(100*sd(myData[,1]),1)
round(100*sd(myData[,5]),1)

```

Try it!

